Popularity, Mixed Matchings, and Self-duality

### 2017-04-03 CO750-ApproximationAlgorithmsTopics/assignments/assignment_03.tex
Wrote up 3b

$$c(R^*)=2c(L^*)$$

@@

Then, since $R$ is a minimum weight $out$-arborescence on $G$ with $c(\bar{L}) = c(R)$ this will suffice to prove the approximation factor as $c(R) \leq c(R^*) = 2c(L^*)$.

@@

Now we change $(WTAP)$ so that the requirement is $(V, E\cup \bar{L})$ is $2$-vertex-connected. Since links between adjacent vertices do not help $2$-vertex-connectivity we may assume there are no such links, otherwise just discard them before starting the algorithm. We will modify the algorithm $\cA$ as follows, the big changes are that we take a leaf to be the root, and we ``shorten" the downward arcs, and we add ``cross" arcs:

@@

@@

>>>>

$$c(R^*)\leq 2c(L^*)$$

@@

Then, since $R$ is a minimum weight $out$-arborescence on $G$ with $c(\bar{L}) = c(R)$ this will suffice to prove the approximation factor as $c(R) \leq c(R^*) \leq 2c(L^*)$.

@@

Now we change $(WTAP)$ so that the requirement is $(V, E\cup \bar{L})$ is $2$-vertex-connected. We may assume we are not in the degenerate case where we have a graph consisting of only two vertices connected by an edge. Such a tree is $2$-vertex-connected by default. We will modify the algorithm $\cA$ as follows, the big changes are that we take a leaf to be the root, and we ``shorten" the downward arcs, and we add ``cross" arcs:

@@

By lemma \ref{lemma:3b-poly} it remains to verify the approximation factor. Let $L^* \subseteq L$ be an optimal solution to $(WTAP)$ on $G$. Let $D$ denote the directed graph which results from the construction in Steps $1$ and $2$ of $\cA$. We will construct from $L^*$ an arc set $A^*$ of cost
$$c(A^*)\leq 4c(L^*)$$
which contains at $out$-arborescence $R^*$ rooted at $r$ on $D$ with cost $c(R^*) \leq 2c(L^*)$.
Then, since $R$ is a minimum weight $out$-arborescence on $G$ with $c(\bar{L}) = c(R)$ this will suffice to prove the approximation factor as $c(R) \leq c(R^*) \leq 2c(L^*)$.

@@

For each $\ell \in L^*$ add the arcs which correspond to $\ell$ (with respect to Step $2$) to $A^*$. At most four arcs are added so we have $c(A^*) \leq 4c(L^*)$. Suppose we also add to $A^*$ the $0$-cost arcs formed in Step $1$ of $\cA$. We want to show that every vertex is reachable from $r$ by a directed path in $A^*$. Then we can compute an arborescence on $A^*$ and use that to bound the optimal solution.
We claim that for every $r' \in V$, $r'$ can reach, by a directed path using arcs of $A^*$, every vertex in the subtree rooted at $r'$. We proceed by induction on the number of vertices of the subtree rooted at $r'$. In the base case the subtree is of size $1$, and hence $r'$ is a leaf and the claim is immediate. Suppose the claim holds for all subtrees with less vertices that the subtree rooted at $r'$. Let $d$ be a descendant of $r'$. We have two cases: either $r' = r$ or not. In the case $r'\neq r$, the deletion of $r'$ disconnects $G$. Let $u_1, \dots, u_k$ be the children of $r'$. Since $r'$ is not the root $r$, $r'$ has some parent $p$. Let $C_{u_1}, \dots, C_{u_k}, C_p$ denote the respective connected components of $G$ after deleting $r'$. Suppose without loss of generality that $d \in C_{u_1}$. Since $(V, E \cup L^*)$ is $2$-vertex-connected, if we think of $C_{u_1}, \dots, C_{u_k}, C_p$ as vertices, there is a subset of links $T \subseteq L^*$ which forms a spanning tree on $C_{u_1}, \dots, C_{u_k}, C_p$. There exists a path from $C_{u_1}$ to $C_g$  using links of $T$. We may assume without loss generality that this path is of the form $C_{u_1}, C_{u_2}, \dots, C_{u_k}, C_p$ since we may simply ignore connected components which do not join this path (decreasing $k$) and relabelling the children.  
Now, for $i \in [k-1]$ the link between $C_{u_{i}}$ and $C_{u_{i+1}}$ induces a directed arc in $A^*$ from $C_{u_{i+1}}$ to $C_{u_i}$ by Step $2(b)$. Call this arc $t_i$, The link between $C_{u_k}$ and $C_p$ induces a directed arc from an ancestor of $u_k$ to a vertex of $C_{u_k}$ by Step $2(a)$ and $2(c)$. Call this arc $t_k$. Now we describe the directed path to $d$. From $r'$ take the up-arcs of $0$-cost to the head of $t_k$ (note the head of $t_k$ may be $r'$, in which case take an empty path). Then follow $t_k$ to a vertex of $C_{u_k}$. We use the following process to traverse $C_k$: Use the $0$-cost arcs to reach $u_k$. Use the inductive hypothesis to reach the head of $t_{k-1}$, and follow $t_{k-1}$ to a vertex of $C_{k-1}$. Repeat this process for each $C_i$ from $C_{k-1}$ to $C_1$. After doing so the directed path has reached the tail of $t_1$. Again follow the $0$-cost arcs up to $u_1$, and use induction to find a directed path from $u_1$ to $d$. All together this process yields a directed path from $r'$ to $d$.
In the induction step the case where $r' =r$ is simpler. Let $u$ be the child of $r'$. Then $d$ is a descendant of $u$ or equal to $u$. In either case, since $L^*$ is feasible there exists some link from $r$ to a descendant of $u$. This link induces an arc from $r$ to a descendant $d'$ of $u$ in $A^*$. Since $r$ is a leaf, the head of this arc is not-moved from $r$ in step $2(c)$. To reach $d$ from $r$, simply traverse the arc from $r$ to $d'$ then traverse the $0$-cost path from $d'$ up to $u$ and use the inductive hypothesis to find a directed path from $u$ to $d$. The resulting concatenated directed path connects $r$ to $d$. Hence by induction, the root $r$ is connected to every vertex of $D$ by a directed path using only arcs of $A^*$.
Therefore we can form a arborescence on $D$, say $R^*$ which is a subset of $A^*$. Choose the minimum cost such $R^*$. We claim that $R^*$ uses at most two arcs corresponding to any given link $uv \in L^*$. Suppose for a contradiction $R^*$ uses at least three arcs corresponding to $uv$. If $R^*$ uses both $(u,v)$ and $(v,u)$ then those two arcs form a directed cycle in $R^*$, contradicting that $R^*$ is acyclic. Hence $R^*$ uses one of $(u,v)$ or $(v,u)$, as well as the two downward arcs formed. Say without loss of generality $R^*$ uses $(u,v)$. Let $p$ be the lowest common ancestor of $u$ and $v$. Let $p_u$ and $p_v$ be the respective children of $p$ on the path to $u$ and $v$ respectively. So $R^*$ uses the arcs $(p_u,u)$ and $(p_v,v)$ as well as $(u,v)$ since $R^*$ uses at least three corresponding arcs. But then $R^*$ has two directed paths from $r$ to $v$. The first from $r$ to $p_u$, and then along $(p_u,u)$ and $(u,v)$ to reach $v$. The second along the directed path from $r$ to $p_v$ then along $(p_v, v)$ to reach $v$. So arc $(u,v)$ is not needed for connectivity in $R^*$, contradicting the optimality of $R^*$.
Therefore we have formed an arborescence $R^*$ such that $c(R^*) \leq 2c(L^*)$. Since $R$ is a minimum cost arborescence and $c(\bar{L}) = c(R)$ we have
$$c(\bar{L}) = c(R) \leq c(R^*) \leq 2c(L^*).$$
So the approximation factor holds as desired.
